---
title: "R Notebook"
output: html_notebook
---
Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r}
library(tidyverse)
```
```{r}
library(tsibble)
library(tsibbledata)
library(lubridate)
library(ggplot2)
library(ggfortify)
library(feasts)
library(fable)
```

```{r}
data <- readxl::read_excel("C:/Users/ddsdi/Desktop/Msc Group Project/final Datset.xlsx")  #R Dataset
data <- data %>%
   mutate(DATEINDEX = yearweek(DATE)) %>%
   as_tsibble(index = DATEINDEX, key = c(SALE_TYPE,VEGETABLE
))                                                             
data
```
```{r}
dataset_VEGETABLE <- data %>%
  filter(VEGETABLE == "CABBAGES")

autoplot(dataset_VEGETABLE, PRICE) +
  labs(title = "CARROTS - Price",
       subtitle = "Dambulla ",
       y = "Price (Rs. per Kg)")
```
```{r}
dataset_VEGETABLE %>%
  gg_season(PRICE, period = "year") +
  labs(y = "Price for Kg",
       title = "Seasonal plot: Price")
```

```{r fig.width=32,fig.height=16}
# fig.width=24,fig.height=16

subuqery_DS <- dataset_VEGETABLE %>% filter(SALE_TYPE == "R") %>% index_by(QUARTER_INDEX = yearmonth(DATE)) %>% group_by_key() %>%
  summarise(
    MEAN_PRICE = mean(PRICE)
  )
  
subuqery_DS %>%
  gg_subseries(MEAN_PRICE) +
  labs(
    y = "Price for Kg",
    x = "Year",
    title = "Sub Series plot: Quarterly Price"
  )

```
```{r}
dataset_VEGETABLE %>% filter(year(DATE)=="2018") %>% autoplot(PRICE)
```
```{r}
dataset_VEGETABLE %>%
  group_by(DATE) %>%
  filter(SALE_TYPE=="R") ->x
dataset_VEGETABLE %>%
  group_by(DATE) %>%
  filter(SALE_TYPE=="W") ->y


combined <- merge(x, y, by.x = "DATE", by.y = "DATE")
combined <- combined %>% select(DATE, PRICE.x, PRICE.y) 
colnames(combined)[colnames(combined) == 'PRICE.x'] <- 'RETAIL'
colnames(combined)[colnames(combined) == 'PRICE.y'] <- 'WHOLESALE'



combined %>%
  #filter(year(DATE) == 2018) %>%
  ggplot(aes(x = RETAIL, y = WHOLESALE)) +
  geom_point() +
  labs(x = "RETAIL",
       y = "WHOLESALE",
       title = "PRICES OF BEANS")


```


```{r}
tempData <- dataset_VEGETABLE %>% 
            filter(SALE_TYPE=="R") %>%
            index_by(INDEX = yearmonth(DATE)) %>% 
            group_by(INDEX) %>%
            summarise(
              MEAN_PRICE = mean(PRICE)
            )


tempData %>% #filter(year(INDEX) == 2018) %>%
  gg_lag(MEAN_PRICE, geom = "point") +
  labs(x = "Mean price for Beans in quarters")
```
```{r}
tempData %>% ACF(MEAN_PRICE, lag_max = 9)
```
```{r}
tempData %>%
  ACF(MEAN_PRICE) %>%
  autoplot() + labs(title="Bean Mean Prices")
```
This plot shows us that data is comparable to "WHITE NOISE" as autocorrelation is close to zero and within the boundaries.

Also when considering the prices it is important to understand the inflation of money as well. But since we dont have a price index, we will not consider that.

```{r}
lambda <- tempData %>%
  features(MEAN_PRICE, features = guerrero) %>%
  pull(lambda_guerrero)

tempData %>%
  autoplot(box_cox(MEAN_PRICE, lambda)) +
  labs(y = lambda,
       title = "Mean Price" ,
         round(lambda,2))
```
```{r}
dataset_VEGETABLE %>% filter(SALE_TYPE=="R") %>%
  model(
    classical_decomposition(PRICE, type = "multiplicative")
  ) %>%
  components() %>%
  autoplot() +
  labs(title = "Classical multiplicative decomposition of MEAN_PRICE for Beans - Retail")
```
We can clearly see that the seasonal decomposition include a recognizable pattern in roughly each 3 months which suggest the influence of various weather and plantation times Sri Lanka uses. (for Beans)

```{r}
x11_dcmp <- tempData %>%
  model(x11 = X_13ARIMA_SEATS(MEAN_PRICE ~ x11())) %>%
  components()
autoplot(x11_dcmp) +
  labs(title =
    "Decomposition of MEAN PRICE of Beans using X-11.")

```
```{r}
x11_dcmp %>%
  ggplot(aes(x = INDEX)) +
  geom_line(aes(y = MEAN_PRICE, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(y = "Price",
       title = " ") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )
```

```{r}
seats_dcmp <- tempData %>%
  model(seats = X_13ARIMA_SEATS(MEAN_PRICE ~ seats())) %>%
  components()
autoplot(seats_dcmp) +
  labs(title = "Decomposition of MEAN_PRICE of Beans using SEATS")
```
```{r}
dataset_VEGETABLE %>% filter(SALE_TYPE=="W") %>%
  model(
    STL(PRICE ~ trend(window = 7) +
                   season(window = 'periodic'),
    robust = TRUE)) %>%
  components() %>%
  autoplot()
```

```{r}

fit <- dataset_VEGETABLE %>% 
            filter(SALE_TYPE=="R") %>%
            filter_index("2018 W01" ~ "2020 W52") %>%
            model( Mean = MEAN(PRICE),
                  `Naïve` = NAIVE(PRICE),
                  `Seasonal naïve` = SNAIVE(PRICE~lag("1 months")))
```

```{r}
fit %>%
  forecast(h = 26) %>%
  autoplot(dataset_VEGETABLE) +
  labs(y = "Price", title = "Price Variation of using Simple forecasting methods")
```

```{r}

fit_dcmp <- dataset_VEGETABLE %>%
  model(stlf = decomposition_model(
    STL(PRICE ~ trend(window = 7), robust = TRUE),
    NAIVE(season_adjust)
  ))

fit_dcmp %>% filter(SALE_TYPE == "R") %>% gg_tsresiduals()
fit_dcmp %>% filter(SALE_TYPE == "W") %>% gg_tsresiduals()

```
```{r}


fit_dcmp %>%
  forecast(h = 10) %>%
  autoplot(dataset_VEGETABLE)+
  labs(y = "Price value",
       title = "Carrot Price forecast using STL decomposition")


```
```{r}
fit_vegetable <- dataset_VEGETABLE %>%
  model(TSLM(PRICE ~ trend() + season()))

fit_vegetable %>% filter(SALE_TYPE=="R") %>% report()
```
```{r}
fit_vegetable %>% filter(SALE_TYPE=="R") %>% augment() %>%
  ggplot(aes(x = DATEINDEX)) +
  geom_line(aes(y = PRICE, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  scale_colour_manual(
    values = c(Data = "black", Fitted = "#D55E00")
  ) +
  labs(y = "Price",
       title = "Mean weekly price of beans") +
  guides(colour = guide_legend(title = "Series"))
```
```{r}
recent_prices <- dataset_VEGETABLE %>%filter_index("2018 W01" ~ "2020 W52") %>%
  filter(SALE_TYPE=="R")

fit_price <- recent_prices %>%
  model(TSLM(PRICE ~ trend() + season()))

fc_price <- forecast(fit_price, h = 26)

fc_price %>%
  autoplot(dataset_VEGETABLE) +
  labs(
    title = "Forecasts of Tomato Price using regression",
    y = "Price"
  )
```
```{r}
# Estimate parameters
fit_SES <- dataset_VEGETABLE %>%filter_index("2018 W01" ~ "2020 W52") %>% filter(SALE_TYPE=="R") %>%
  model(ETS(PRICE ~ error("A") + trend("N") + season("N")))

fc_SES <- fit_SES %>%
  forecast(h = 26)

fc_SES %>%
  autoplot(dataset_VEGETABLE %>% filter(SALE_TYPE=="R")) +
  geom_line(aes(y = .fitted), col="#A55E00",
            data = augment(fit)) +
  labs(y="PRICE", title="Tomato Price variation") +
  guides(colour = "none")

report(fit_SES)
```

```{r}

fit_with_seasonality <- dataset_VEGETABLE%>%filter_index("2018 W01" ~ "2020 W52") %>%
  model(
    additive = ETS(PRICE ~ error("A") + trend("A") + season("A", period=4)),
    multiplicative = ETS(PRICE ~ error("M") + trend("A") + season("M", period=4))
  )

fc_with_seasonality <- fit_with_seasonality %>% forecast(h = 26)

fc_with_seasonality %>%
  autoplot(dataset_VEGETABLE) +
  labs(title="Cabbage Prices",
       y="Price") +
  guides(colour = guide_legend(title = "Forecast"))

report(fit_with_seasonality %>% filter(SALE_TYPE=="R") %>% select("additive"))

```
```{r}
components(fit_with_seasonality) %>%
  autoplot() +
  labs(title = "ETS components")
```
```{r}
dataset_VEGETABLE %>%
  filter(SALE_TYPE == "W") %>%
  transmute(
    `Log Prices` = log(PRICE),
    `Single differenced log prices` = difference(log(PRICE), 1),
    `Double differenced log prices` =
                     difference(difference(log(PRICE), 1), 1)
  ) %>%
  pivot_longer(-DATEINDEX, names_to="Type", values_to="Price") %>%
  mutate(
    Type = factor(Type, levels = c(
      "Log Prices",
      "Single differenced log prices",
      "Double differenced log prices"))
  ) %>%
  ggplot(aes(x = DATEINDEX, y = Price)) +
  geom_line() +
  facet_grid(vars(Type), scales = "free_y") +
  labs(title = "Carrot Prices - Wholesale", y = NULL)


```
```{r}
dataset_VEGETABLE%>%
  features(PRICE, unitroot_kpss)

dataset_VEGETABLE%>%
  features(PRICE, unitroot_ndiffs)
```
Above values indicate that the series is stationary, even without doing any data transformations. therefore no need to differencing. same goes to seasonal differencing as well.

```{r}
dataset_VEGETABLE %>%
  features(PRICE, unitroot_nsdiffs)
```
```{r}
fit <- dataset_VEGETABLE%>%filter_index("2018 W01" ~ "2020 W52") %>%
  model(ARIMA(PRICE))
report(fit%>% filter(SALE_TYPE=="R"))
```

```{r}
fit %>% forecast(h=26) %>%
  autoplot(dataset_VEGETABLE) +
  labs(y = "Price", title = "Tomato Prices")
```

```{r}
fit2 <- dataset_VEGETABLE %>%
  model(ARIMA(PRICE ~ pdq(1,0,0)))
report(fit2 %>% filter(SALE_TYPE=="R"))

fit2 %>% forecast(h=20) %>%
  autoplot(dataset_VEGETABLE) +
  labs(y = "Price", title = "Beans Prices")
```

```{r}

```



